### 2019-07-01 README.Rmd
Spell Check and knit draft of README for github

The goal of diffEnrich is simple. Given a list of gene symbols, *diffEnrich*  will
perform differential enrichment ananlysis using the Kyoto Encyclopedia of Genes
and Genomes (KEGG) REST API. This package provides a number of functions that are 
intended to be used in a pipeline (See Figure N). Breifly, the workflow will
download and clean species specific ENTREZ gene IDs and map them to their respective
KEGG pathways by accessing KEGG's REST API. This way the user will always have 
the most up-to-date pathway data from KEGG. Next, the user will identify significantly
enriched pathways from two different gene sets, and finally, the user will identify 
pathways that are differentially expressed between the two gene sets. In addition to 
the analysis pipeline, this package also provides functions for data visualizations.

@@

### Step 1: Collect and clean pathways from KEGG API
First we will use the *get_kegg* function access the KEGG REST API and download the data sets required to perform 
our downstream analysis. This function takes two arguments. The first, 'species' is required and is the species of 
interest. Currently, *diffEnrich* supports three species, and the argument is a character string using the KEGG
code [REF](https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf): 
Homo sapiens (human), use 'hsa'; Mus musculus (mouse), use 'mmu', and Rattus norvegicus (rat), use 'rno'. The second,
'path' is also passed as a character string, and is the path of the directory in which the user would like to
write the data sets downloaded from the KEGG REST API. If the user does not provide a path, the data sets will 
be automatically written to the current working directory using the *here::here()* functionality. These data 
sets will be tab delimited files with a name describing the data, and for reproducibility, the date they were generated
and the version of KEGG when the API was accessed. In addition to these flat files, *get_kegg* will also
create a named list with the three relavent KEGG data sets. The names of this list will describe the the data set.

@@

In this step we will use the *pathEnrich* function to identify KEGG pathways that are enriched (i.e. over-represented) based on a list of genes we
are interested in and based on a list of background genes. This function may not always use the complete list of genes provided 
by the user. Specifically, it will only use the genes from the list provided that are also in
the most current species list pulled from the KEGG REST API using *get_kegg*, or from the older KEGG data loaded by the user from a 
previous *get_kegg* call. The *pathEnrich* function should be run at least twice, once for the genes of interest and once for the background.
Each *pathEnrich* call generates a dataframe summarizing the results of an enrichment analysis in which a Fisher's
Exact test is used to identify which KEGG pathways are enriched by the user's list of interesting genes with repect to background
enrichment, and a p-value is calculated using a hypergeometric distribution (Formula 1). P-values are adjusted for multiple comparisons by controlling the False Discovery Rate (FDR) at 0.05.

@@

*pathEnrich* generates a dataframe with 9 columns described below.

@@

>>>>

The goal of diffEnrich is simple. Given a list of gene symbols, *diffEnrich*  will
perform differential enrichment analysis using the Kyoto Encyclopedia of Genes
and Genomes (KEGG) REST API. This package provides a number of functions that are 
intended to be used in a pipeline (See Figure N). Briefly, the workflow will
download and clean species specific ENTREZ gene IDs and map them to their respective
KEGG pathways by accessing KEGG's REST API. This way the user will always have 
the most up-to-date pathway data from KEGG. Next, the user will identify significantly
enriched pathways from two different gene sets, and finally, the user will identify 
pathways that are differentially expressed between the two gene sets. In addition to 
the analysis pipeline, this package also provides functions for data visualizations.

@@

### Step 1: Collect and clean pathways from KEGG API
First we will use the *get_kegg* function access the KEGG REST API and download the data sets required to perform 
our downstream analysis. This function takes two arguments. The first, 'species' is required and is the species of 
interest. Currently, *diffEnrich* supports three species, and the argument is a character string using the KEGG
code [REF](https://www.pnas.org/content/suppl/2008/09/11/0806162105.DCSupplemental/ST1_PDF.pdf): 
Homo sapiens (human), use 'hsa'; Mus musculus (mouse), use 'mmu', and Rattus norvegicus (rat), use 'rno'. The second,
'path' is also passed as a character string, and is the path of the directory in which the user would like to
write the data sets downloaded from the KEGG REST API. If the user does not provide a path, the data sets will 
be automatically written to the current working directory using the *here::here()* functionality. These data 
sets will be tab delimited files with a name describing the data, and for reproducibility, the date they were generated
and the version of KEGG when the API was accessed. In addition to these flat files, *get_kegg* will also
create a named list with the three relevant KEGG data sets. The names of this list will describe the the data set.

@@

In this step we will use the *pathEnrich* function to identify KEGG pathways that are enriched (i.e. over-represented) based on a list of genes we
are interested in and based on a list of background genes. This function may not always use the complete list of genes provided 
by the user. Specifically, it will only use the genes from the list provided that are also in
the most current species list pulled from the KEGG REST API using *get_kegg*, or from the older KEGG data loaded by the user from a 
previous *get_kegg* call. The *pathEnrich* function should be run at least twice, once for the genes of interest and once for the background.
Each *pathEnrich* call generates a data frame summarizing the results of an enrichment analysis in which a Fisher's
Exact test is used to identify which KEGG pathways are enriched by the user's list of interesting genes with respect to background
enrichment, and a p-value is calculated using a hypergeometric distribution (Formula 1). P-values are adjusted for multiple comparisons by controlling the False Discovery Rate (FDR) at 0.05.

@@

*pathEnrich* generates a data frame with 9 columns described below.

@@

### Step 3: Identify differentially enriched KEGG pathways

The diffEnrich function will merge the results from the *pathEnrich* calls generated above. Specifically, the data frame
'sig_pe' and the data frame 'bkg_pe' will be merged using the base *merge* functionality and be merged by the following columns:
'c("KEGG_ID", "KEGG_description", "KEGG_cnt", "numTested")'. This merged data set will then be used to perform differential 
enrichment using the same method and p-value calculation as described above.

```{r diffEnrich}
## Perform differential enrichment
diff_enrich <- diffEnrich(sig_pe = sig_pe, bkg_pe = bkg_pe)

knitr::kable(head(diff_enrich),
             caption = "Table 2. Description of columns is dataframe generated by pathEnrich", row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```

The result of the *diffEnrich* call is the merged data frame with the estimated odds ration generated by the Fisher's Exact test 
and the associated p-value.
